### 2015-04-26 draft.tex
Progress on the hardware and design decisions.

\documentclass[letterpaper, oneside, 10pt]{report}
\usepackage[margin = 1.0in]{geometry}
\usepackage[nottoc,notlof,notlot]{tocbibind}
\usepackage[hidelinks]{hyperref}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{setspace}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{cite}
\usepackage{classicthesis}

\geometry{letterpaper}

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}

\setlength{\skip\footins}{0.8cm}
\setlength{\tabcolsep}{30pt}

% Title page parameters
\lhead{Andrew Lawson}
\title{\large COOPERATIVE 3-D MAP GENERATION USING MULTIPLE UAVS}
\author{
  \large
  Andrew Lawson \\
  University of Connecticut \\
  Computer Science and Engineering
}
\date{\large May 1, 2015}
\pagestyle{fancy}

\begin{document}
% \doublespacing

% Title page
\begin{titlepage}
  \maketitle
  \thispagestyle{empty}
\end{titlepage}
\clearpage

% Abstract
\begin{abstract}
  In the modern field of robotics, the ability for a robot to smoothly move about the environment, regardless of the medium, has proven to be fairly trivial. However, the ability for a mobile robot to track its own position and make deterministic decisions about it's own location in a foreign environment remains a fundamental, but critical problem. In most typical robotic experiments, the use of external measurement systems (such as GPS or motion capture cameras) allow researchers to capture the exact position of a robot in its environment and focus on other areas of interest. In the past couple decades, successful algorithms called SLAM\footnote{Simultaneous-Localization-And-Mapping: algorithm that uses sensor data to independently determine a position relational to an environment and simultaneously create a virtual map} have allowed singular mobile robots to determine their own position in a world space and construct a 3-D map from collected positional data. However, despite these advances, the ability for multiple UAV\footnote{UAV: unmanned aerial vehicle} robots to cooperatively determine their environmental boundaries is an area that remains relatively under-developed. Some studies have attempted this decentralized, cooperative approach, but fall short either by depending on external systems or lacking true 3-D capability. By taking this distributional approach with multiple robotic agents, the computational overhead of map generation is greatly reduced. Teams of coordinated robots could efficiently explore unaccessible areas, complete team-based tasks (like lifting large objects), conduct dangerous search and rescue missions, and ultimately increase the feasibility of robotics in numerous situations.
\end{abstract}
\clearpage

\renewcommand{\abstractname}{Acknowledgements}
\begin{abstract}
 I'd like to thank Prof. Shalabh Gupta, Prof. Ashwin Dani, and Prof. Robert McCartney for their advice and support. I'd also like to thank Kris Balisciano, Chung Yang, and everyone in L.I.N.K.S. Lab for all their work on the project.
\end{abstract}
\clearpage

% Table of Contents, Figures, and Tables
\tableofcontents
\listoffigures
\listoftables
\clearpage

\chapter{Introduction}

\section{Overview}
\section{Literature Review}

In recent research projects with quadcopters, a wide variety of methods have been used to localize a robot in their environment, including (but not limited to) the use of high-resolution laser scanners \cite{bachrach2011range,shen2011autonomous}, monocular cameras \cite{soundararaj2009autonomous}, and ``RGB-D"\footnote{RGB-D: red-blue-green-depth} cameras as a combination of the two \cite{huang2011visual}. A laser has been used to generate a point cloud and pull out contours that trace the environment\cite{bachrach2011range}, a monocular camera to exploit texture gradients and continuations that can translated into depth estimates \cite{soundararaj2009autonomous}, and Microsoft Kinects (RGB-D cameras) to associate laser depth with image data \cite{huang2011visual}. Lasers capable of such precision and frequency are economically infeasible, and additionally, a pure visual approach is slow, can result in large margins of error, and lacks precision. Using a combination of the two, by means of an RGB-D camera, is responsive,  inexpensive, and produces results of reasonable reliability.

With previous RGB-D camera approaches, there have been five primary components: feature extraction, rotation estimation, feature matching, inlier detection, and motion estimation \cite{huang2011visual}. This involves taking certain features (like color, edges, corners, etc.) out of an image frame, estimating the rotation pose of the robot, matching new features to existing ones, filtering out bad data, and then estimating the motion of the robot. Using a data filter called an Extended-Kalman-Filter, the visual odometry data will be combined with velocity estimates to form a state estimation of the robot \cite{huang2011visual}. By implementing this filter, this will allow real-time positional estimates of the vehicle.

Researchers have also attempted de-centralized SLAM methods like I have proposed, but do not provide total three-dimensional capability. Some have used a VICON camera system \cite{leung2012decentralized} to provide true location data to robotic agents. Others have implemented truly autonomous, cooperative SLAM, but only focused on 2-D, planar data analysis \cite{cunningham2012large}. As I have mentioned previously, this does not fully address the ability to move and coordinate multiples robot entities independently in a realistic world setting. By examining this precedent work in the field, however, the ability to create cooperative 3-D SLAM can be fully visualized.

\chapter{Robot Design}

\section{Design Decisions}

Several key design decisions arose when choosing a direction for robot design and selection of components. The most signifcant choice for the robot was choosing an architecture for the onboard computational unit - specifically an ARM or x86 computer. The second key design dilemma involved the selection of a laser-scanning sensor or an RGB-D camera. As shown below, both these choices had great effect upon the path of experimentation and results obtained.

\subsection{ARM vs. x86}

The main advantage of choosing an x86 board (found in most desktop or laptop computers) is compatibility. Almost all software packages are available in prebuilt, binary form on package managers (such as Ubuntu's apt-get) and setting up the computing environment can be done with a few simple installation commands. This convenience, unfortunately, comes at a financial and dimensional price - most small form-factor x86 boards are extremely expensive and cheaper options are considerably larger in size. These tradeoffs make selection of such a board unfeasible for an aerial drone where space is extremely limited and minimization of cost is a primary concern. As a result, ARM boards (like those in smartphones) are much more attractive - they offer extremely small form-factors, yet remain very affordable. The issue that arises, of course, is that of compatibility - many programs or libraries are not (immediately) available on an ARM installation and require manual compilation of source code using an ARM-based compiler and possible changes to the source code itself first.

\subsection{Laser vs. RGB-D Camera}

\section{Hardware}

The hardware design of the robot was focused on using low-cost, off-the-shelf components (see Table 1). The approximate cost of a built robot was \$894, significantly cheaper than those used at MIT or University of Washington with similar project requirements.

% Table of components used on robot
\begin{table}[h!]
  \centering
  \caption{Table of Components}
  \vspace{2mm}
  \begin{tabular}{l l l}
    \hline \hline
    \vspace{-2mm}
    Component & \multicolumn{1}{l}{Description} & \multicolumn{1}{l}{Cost} \\ [1ex]
    \hline
    & \\
    DJI F450 Frame & Chassis & \$200 \\
    DJI 1045 Propellers & Props & \$10 \\
    DJI 2212 Motors & Actuators & \$30 \\
    12V and 5V Regulators & Power Reg. & \$40 \\
    Zippy 2450 mAh LiPo & Battery & \$30 \\
    64GB ODROID XU3 & ARM SBU & \$240 \\
    Pixhawk PX4 & Controller & \$220 \\
    Microsoft Kinect & Visual Sensor & \$99 \\
    MaxSonar EZ0	& Sonar Sensor & \$25 \\
  \end{tabular}
\end{table}

\noindent The robot used a stock F450 frame from DJI, which was not modified in any way - all components were mounted on existing areas using screws or adhesive products.

\begin{figure}[h!]
 \caption{Photograph of finished quadcopter robot.}
 \centering
   \includegraphics[width=0.3\textwidth]{figures/quadcopter}
\end{figure}

\section{Software}

The direction of my project has two major stages to complete: Designing the robot and developing the navigation and map-building algorithms. \\

%\begin{figure}[h!]
 %\caption{Prototype quadcopter.}
 %\centering
   %\includegraphics[width=0.6\textwidth]{copter}
%\end{figure}

\chapter{Single Robot}
\section{Navigation}
\section{Mapping}
\section{Results}

\chapter{Multiple Robots}
\section{Map Fusion}
\section{Results}

\chapter{Conclusions}

\chapter{Appendix}

\clearpage

\bibliography{uscholar}
\bibliographystyle{ieeetr}

\end{document}